|

### 2022-06-23 docs/rmd/paper.Rmd
autogit

<!-- Selection --->
The overall performance of GP can depend strongly on the choice of its underlying operators, one crucial component in this is the operator for parent selection. 
Tournament Selection is a commonly used selection operator in EC and is the most used operator in GP systems [@10.1007/978-3-642-16493-4_19, p.181]. A parent solution is selected by randomly sampling $k$ individuals from the current population into a tournament pool and then the solution with the highest fitness score from the tournament pool is selected [@10.1007/978-3-642-16493-4_19, p.182]. 
Lexicase selection has been suggested as an alternative to tournament selection that is not based on aggregating fitness scores. It samples $n$ test cases in random order and then eliminates solutions from the selection pool on a per test case basis if they are not performing on an elite level [@6920034, p.1].

@@

<!-- Generalization --->
An important quality of all supervised machine learning applications, including GP, is the ability to not only optimize performance for the test cases a model is trained on but to also perform well on previously unseen cases, this is referred to as generalization. In most real world applications of symbolic regression only a small subset of labeled data is available for training. The aim is to produce a model that not only accurately predicts the provided training data but can also predict previously unseen cases with high precision [@Gonalves2016AnEO, p.6]. A model that is extensively optimized on the provided training data, can become over fitted to this data sample which may lead to a decrease in generalization.

@@

@@

Both articles argue that many of the problem domains that GP is commonly used for are prevalent for problems that are modal/ uncompromising in nature. In comparison to other selection operators, populations that are evolved using variations of the lexicase selection operator show a very high degree of genetic diversity which might be a key contributor to the improved performance [@6920034, p.1] [@epsilon_lexicase_main, p.745]. In theory, Lexicase based selection should be more prone to select solutions that are specialist (solutions that achieve high fitness on a smaller subset of unusual training cases) than solutions that are generalist (solutions that achieve high average fitness among all training cases).

@@

More recent research by @epsilon_lexicase_main suggests $\epsilon$-lexicase selection as a modified selection operator that can improve overall GP performance if applied to continuous-valued symbolic regression tasks in comparison to tournament and standard lexicase selection.  @epsilon_lexicase_main argues that the original concept behind lexicase selection as formulated by @lexicase_first_desciption does not fit the requirements of real-world symbolic regression tasks. The authors identifies the pass condition of regular lexicase selection as the main problem if applied to symbolic regression: Individual solutions can only be selected if they perform on an elite level but with continuous-valued, and often noisy, data it is very unlikely for two individuals to achieve an exactly equal error for a given test case. This stringent pass condition can result in filtering out too many individuals during the selection process [@epsilon_lexicase_main, p.742]. $\epsilon$-lexicase adresses this problem by introducing an $\epsilon$ parameter that specifies a range around the elite error, individuals that perform inside this range pass the current selection iteration. @https://doi.org/10.48550/arxiv.1709.05394 describes $\epsilon$-lexicase selection as a more "relaxed version of lexicase selection".
Different methods to configure the $\epsilon$ Parameter have been explored, for the limited scope of this project I will focus on the most promising implementation of $\epsilon$-lexicase selection that is automatically adjusted on the basis of the median absolute deviation of errors inside the selection pool [@epsilon_lexicase_main, p.742] [@https://doi.org/10.48550/arxiv.1709.05394, p.6].

@@

1. Sample:
  (a) Randomly sample k individuals from the current population
      into a tournament pool.
2. Select:
  (a) Compute the mean fitness for each individual inside the tournament pool
      based on all fitness cases.
  (b) Select individual from the tournament pool with the highest mean fitness.
```

@@

Generalization, the ability of a model to perform well on previously unseen cases, is one of the fundamental goals in most real world machine learning applications. @open_issues_gp raises awareness to the fact, that the topic of generalization in GP has not gotten the attention that other machine learning-based areas  have attributed to it. A similar statement has been proposed by @generalisation_in_gp in his review of research in generalization in GP. Both authors address the need for additional research regarding the topic.

@@

An aspect that is suspected to negatively correlate with GP's generalization is overall size and in particular bloating of the resulting programs. Bloating can be described as a growth in the total size of a program that does not also improve it's performance in any meaningful way [@bloat_overfitting_gp, p. 1].
It has been widely suspected that GP configurations that produce very large programs also have a higher tendency to specialize on difficult or unusual test cases which could result in lower generalization [@wang_wagner_rondinelli_2019, p. 268]. The minimum description length principle (MDLP) describes this phenomenon by claiming that less complex models are more likely to perform better at generalizing than more complex models that achieve a comparable fitness during the training phase [@open_issues_gp, p. 349]. However MDLP should be considered carefully, experiments by @bloat_overfitting_gp on the prediction of bio availability for drugs demonstrate that GP based techniques can either bloat without over fitting or over fit without bloating [@bloat_overfitting_gp, p. 8].

@@

Since it's inception, GP based symbolic regression has been widely used in practical applications ranging over a wide field of disciplines. A well known example of an industrial use-case has been published by @soft_sensor_gp. The authors use GP based symbolic regression to derive nonlinear functions that can be used to significantly increase the robustness of industrial sensors which resulted in wide-spread adoption and a significant reduction in costs.

@@

Using two otherwise identical GP systems, one deploying tournament selection and the other $\epsilon$-lexicase selection, the objective is to find a computer program that best predicts the outcome variable heating load $(Y1)$^[The symbolic regression is performed on one of the two provided outcome variables, the variable Cooling Load will be excluded.] of the buildings using a subset of the eight building attributes $(X1,..,X8)$ for it's input. The specific meaning of all attributes inside the data set are described in table 1.

@@

The statistical analysis of the collected data will first focus on examining the question if, on average, the usage of $\epsilon$-lexicase selection will result in models that perform significantly different than models that are evolved using tournament selection both.

@@

## Gentetic Programming Configuration

@@

The primitive set consists of the terminals that are listed in table 3 and the functions that are listed in table 4. To avoid run time errors I implemented protected version of the operators for division, natural logarithm and square root [@koza_main, p.82-83]. As suggested by @koza_main I also included ephemeral constants to the set of terminals to provide the evolutionary search the opportunity to explore and include randomly generated constants.

@@

To gain further insight into the differences in generalization behavior and to address the research question with a higher level of certainty, the experiment I conducted should be repeated for different GP configurations. Some exemplary parameters that could drastically influence the results of this experiment include the population size, the total number of generations, the inclusion of elitism or the computation of $\epsilon$ in lexicase selection. Other factors whose influence on generalization behavior could be important include different methods to compute an individuals fitness (e.g. mean absolute error or root mean squared error), testing different variations of the genetic operators for mutation and crossover and experimenting with different variations of the training:testing split size.

Another obvious weakness of my project is that the experiment is only based on a single symbolic regression application, the prediction of the heating load of buildings based on a relatively small data set. To gain further trust in the obtained results, the experiment should be repeated for other symbolic regression tasks and data sets. At the current state it remains unknown if the generalization behavior of both selection operators might be highly problem specific and dependent on the data set used for training and testing.

>>>>

<!-- Selection --->
The overall performance of GP can depend strongly on the choice of its underlying operators, one crucial component in this is the operator for parent selection. 
Tournament Selection is a commonly used selection operator in EC and is the most used operator in GP systems [@10.1007/978-3-642-16493-4_19, p.181]. A parent solution is selected by randomly sampling $k$ individuals from the current population into a tournament pool and then the solution with the highest fitness score from the pool is selected [@10.1007/978-3-642-16493-4_19, p.182]. 
Lexicase selection has been suggested as an alternative to tournament selection that is not based on aggregating fitness scores. It samples $n$ test cases in random order and then eliminates solutions from the selection pool on a per test case basis if they are not performing on an elite level [@6920034, p.1].

@@

<!-- Generalization --->
An important quality of all supervised machine learning applications, including GP, is the ability to not only optimize performance for the test cases a model is trained on but to also perform well on previously unseen cases, this is referred to as generalization. In most real world applications of symbolic regression only a small subset of labeled data is available for training. The aim is to produce a model that not only accurately predicts the provided training data but can also predict previously unseen cases with high precision [@Gonalves2016AnEO, p.6]. A model that is extensively optimized on training data, can become over fitted to the data sample which may lead to a decrease in generalization.

@@

<!-- Vorgehen --->
Section 2 will give a brief summary and discussion of topic related scientific research. Section 3 describes the design of the experiment as well as GP specific implementation details. The statistical  evaluation of all relevant results and the summarized conclusions are given in section 3 and 4. Section 5 addresses limitation and open questions of this research project.

@@

Both articles argue that many of the problem domains that GP is commonly used for are prevalent for problems that are modal or uncompromising in nature. In comparison to other selection operators, populations that are evolved using variations of the lexicase selection operator show a very high degree of genetic diversity which might be a key contributor to the improved performance [@6920034, p.1] [@epsilon_lexicase_main, p.745]. In theory, Lexicase based selection should be more prone to select solutions that are specialist (solutions that achieve high fitness on a smaller subset of unusual training cases) than solutions that are generalist (solutions that achieve high average fitness among all training cases).

@@

More recent research by @epsilon_lexicase_main suggests $\epsilon$-lexicase selection as a modified selection operator that can improve overall GP performance if applied to continuous-valued symbolic regression tasks in comparison to tournament and standard lexicase selection.  @epsilon_lexicase_main argues that the original concept behind lexicase selection as formulated by @lexicase_first_desciption does not fit the requirements of real-world symbolic regression tasks. The authors identifies the pass condition of regular lexicase selection as the main problem if applied to symbolic regression: Individual solutions can only be selected if they perform on an elite level but with continuous-valued, and often noisy, data it is very unlikely for two individuals to achieve an exact equal error for a given test case. This stringent pass condition can result in filtering out too many individuals during the selection process [@epsilon_lexicase_main, p.742]. $\epsilon$-lexicase adresses this problem by introducing an $\epsilon$ parameter that specifies a range around the elite error, individuals that perform inside this range pass the current selection iteration. @https://doi.org/10.48550/arxiv.1709.05394 describes $\epsilon$-lexicase selection as a more "relaxed version of lexicase selection".
Different methods to configure the $\epsilon$ Parameter have been explored, for the limited scope of this project I will focus on the most promising implementation of $\epsilon$-lexicase selection that is automatically adjusted on the basis of the median absolute deviation of errors inside the selection pool [@epsilon_lexicase_main, p.742] [@https://doi.org/10.48550/arxiv.1709.05394, p.6].

@@

1. Sample:
  (a) Randomly sample k individuals from the current population
      into a tournament pool.
2. Select:
  (a) Compute the mean fitness for each individual inside the tournament pool
      based on all fitness cases.
  (b) Select individual from the tournament pool with the highest mean fitness score.
```

@@

Generalization, the ability of a model to perform well on previously unseen cases, is one of the fundamental goals in most real world machine learning applications. @open_issues_gp raises awareness to the fact, that the topic of generalization in GP has not gotten the attention that other machine learning-based areas attributed to it. A similar statement has been proposed by @generalisation_in_gp in his review of research in generalization in GP. Both authors address the need for additional research regarding the topic.

@@

An aspect that is suspected to negatively correlate with GP's generalization is overall size and in particular bloating of the resulting programs. Bloating can be described as a growth in the total size of a program that does not improve its performance in any meaningful way [@bloat_overfitting_gp, p. 1].
It has been widely suspected that GP configurations that produce very large programs also have a higher tendency to specialize on difficult or unusual test cases which could result in lower generalization [@wang_wagner_rondinelli_2019, p. 268]. The minimum description length principle (MDLP) describes this phenomenon by claiming that less complex models are more likely to perform better at generalizing than more complex models that achieve a comparable fitness during the training phase [@open_issues_gp, p. 349]. However MDLP should be considered with caution, experiments by @bloat_overfitting_gp on the prediction of bio availability for drugs demonstrate that GP based techniques can either bloat without over fitting or over fit without bloating [@bloat_overfitting_gp, p. 8].

@@

Since its inception, GP based symbolic regression has been widely used in practical applications ranging over a wide field of disciplines. A well known example of an industrial use-case has been published by @soft_sensor_gp. The authors use GP based symbolic regression to derive nonlinear functions that can be used to significantly increase the robustness of industrial sensors. This resulted in wide-spread adoption and a significant reduction in costs.

@@

Using two otherwise identical GP systems, one deploying tournament selection and the other $\epsilon$-lexicase selection, the objective is to find a computer program that best predicts the outcome variable heating load $(Y1)$^[The symbolic regression is performed on one of the two provided outcome variables, the variable Cooling Load will be excluded.] of the buildings using a subset of the eight building attributes $(X1,..,X8)$ for its input. The specific meaning of all attributes inside the data set are described in table 1.

@@

The statistical analysis of the collected data will first focus on examining the question if, on average, the usage of $\epsilon$-lexicase selection will result in models that perform significantly different than models that are evolved using tournament selection:

@@

## Genetic Programming Configuration

@@

The primitive set consists of the terminals listed in table 3 and the functions listed in table 4. To avoid run time errors I implemented protected version of the operators for division, natural logarithm and square root [@koza_main, p.82-83]. As suggested by @koza_main I also included ephemeral constants to the set of terminals to provide the evolutionary search the opportunity to explore and include randomly generated constants.

@@

To gain further insight into the differences in generalization behavior and to address the research question with a higher level of certainty, the experiment I conducted should be repeated for different GP configurations. Some exemplary parameters that could drastically influence the results of this experiment include the population size, the total number of generations, the inclusion of elitism or the computation of $\epsilon$ in lexicase selection. Other factors whose influence on generalization behavior could be important include different methods to compute an individuals fitness (e.g. mean absolute error or root mean squared error), testing different variations of the genetic operators for mutation and crossover and experimenting with different variations of the training:testing data set split size.

Another obvious weakness of this project is that the experiment is only based on a single symbolic regression application, the prediction of the heating load of buildings based on a relatively small data set. To gain further trust in the obtained results, the experiment should be repeated for other symbolic regression tasks and data sets. It remains unknown if the generalization behavior of both selection operators might be highly problem specific and dependent on the data set used for training and testing.

### 2022-06-23 docs/rmd/paper.tex
autogit

The overall performance of GP can depend strongly on the choice of its
underlying operators, one crucial component in this is the operator for
parent selection. Tournament Selection is a commonly used selection
operator in EC and is the most used operator in GP systems (Fang and Li,
2010, p. 181). A parent solution is selected by randomly sampling \(k\)
individuals from the current population into a tournament pool and then
the solution with the highest fitness score from the tournament pool is
selected (Fang and Li, 2010, p. 182). Lexicase selection has been
suggested as an alternative to tournament selection that is not based on
aggregating fitness scores. It samples \(n\) test cases in random order
and then eliminates solutions from the selection pool on a per test case
basis if they are not performing on an elite level (Helmuth, Spector and
Matheson, 2015, p. 1).

@@

An important quality of all supervised machine learning applications,
including GP, is the ability to not only optimize performance for the
test cases a model is trained on but to also perform well on previously
unseen cases, this is referred to as generalization. In most real world
applications of symbolic regression only a small subset of labeled data
is available for training. The aim is to produce a model that not only
accurately predicts the provided training data but can also predict
previously unseen cases with high precision (Gonçalves, 2016, p. 6). A
model that is extensively optimized on the provided training data, can
become over fitted to this data sample which may lead to a decrease in
generalization.

@@

@@

Both articles argue that many of the problem domains that GP is commonly
used for are prevalent for problems that are modal/ uncompromising in
nature. In comparison to other selection operators, populations that are
evolved using variations of the lexicase selection operator show a very
high degree of genetic diversity which might be a key contributor to the
improved performance (Helmuth, Spector and Matheson, 2015, p. 1) (La
Cava, Spector and Danai, 2016, p. 745). In theory, Lexicase based
selection should be more prone to select solutions that are specialist
(solutions that achieve high fitness on a smaller subset of unusual
training cases) than solutions that are generalist (solutions that
achieve high average fitness among all training cases).

@@

More recent research by La Cava, Spector and Danai (2016) suggests
\(\epsilon\)-lexicase selection as a modified selection operator that
can improve overall GP performance if applied to continuous-valued
symbolic regression tasks in comparison to tournament and standard
lexicase selection. La Cava, Spector and Danai (2016) argues that the
original concept behind lexicase selection as formulated by Spector
(2012) does not fit the requirements of real-world symbolic regression
tasks. The authors identifies the pass condition of regular lexicase
selection as the main problem if applied to symbolic regression:
Individual solutions can only be selected if they perform on an elite
level but with continuous-valued, and often noisy, data it is very
unlikely for two individuals to achieve an exactly equal error for a
given test case. This stringent pass condition can result in filtering
out too many individuals during the selection process (La Cava, Spector
and Danai, 2016, p. 742). \(\epsilon\)-lexicase adresses this problem by
introducing an \(\epsilon\) parameter that specifies a range around the
elite error, individuals that perform inside this range pass the current
selection iteration. La Cava \emph{et al.} (2017) describes
\(\epsilon\)-lexicase selection as a more ``relaxed version of lexicase
selection''. Different methods to configure the \(\epsilon\) Parameter
have been explored, for the limited scope of this project I will focus
on the most promising implementation of \(\epsilon\)-lexicase selection
that is automatically adjusted on the basis of the median absolute
deviation of errors inside the selection pool (La Cava, Spector and
Danai, 2016, p. 742) (La Cava \emph{et al.}, 2017, p. 6).

@@

1. Sample:
  (a) Randomly sample k individuals from the current population
      into a tournament pool.
2. Select:
  (a) Compute the mean fitness for each individual inside the tournament pool
      based on all fitness cases.
  (b) Select individual from the tournament pool with the highest mean fitness.
\end{verbatim}

@@

Generalization, the ability of a model to perform well on previously
unseen cases, is one of the fundamental goals in most real world machine
learning applications. O'Neill \emph{et al.} (2010) raises awareness to
the fact, that the topic of generalization in GP has not gotten the
attention that other machine learning-based areas have attributed to it.
A similar statement has been proposed by Kushchu (2002) in his review of
research in generalization in GP. Both authors address the need for
additional research regarding the topic.

@@

An aspect that is suspected to negatively correlate with GP's
generalization is overall size and in particular bloating of the
resulting programs. Bloating can be described as a growth in the total
size of a program that does not also improve it's performance in any
meaningful way (Silva and Vanneschi, 2009, p. 1). It has been widely
suspected that GP configurations that produce very large programs also
have a higher tendency to specialize on difficult or unusual test cases
which could result in lower generalization (Wang, Wagner and Rondinelli,
2019, p. 268). The minimum description length principle (MDLP) describes
this phenomenon by claiming that less complex models are more likely to
perform better at generalizing than more complex models that achieve a
comparable fitness during the training phase (O'Neill \emph{et al.},
2010, p. 349). However MDLP should be considered carefully, experiments
by Silva and Vanneschi (2009) on the prediction of bio availability for
drugs demonstrate that GP based techniques can either bloat without over
fitting or over fit without bloating (Silva and Vanneschi, 2009, p. 8).

@@

Since it's inception, GP based symbolic regression has been widely used
in practical applications ranging over a wide field of disciplines. A
well known example of an industrial use-case has been published by
Jordaan \emph{et al.} (2004). The authors use GP based symbolic
regression to derive nonlinear functions that can be used to
significantly increase the robustness of industrial sensors which
resulted in wide-spread adoption and a significant reduction in costs.

@@

Using two otherwise identical GP systems, one deploying tournament
selection and the other \(\epsilon\)-lexicase selection, the objective
is to find a computer program that best predicts the outcome variable
heating load \((Y1)\)\footnote{The symbolic regression is performed on
  one of the two provided outcome variables, the variable Cooling Load
  will be excluded.} of the buildings using a subset of the eight
building attributes \((X1,..,X8)\) for it's input. The specific meaning
of all attributes inside the data set are described in table 1.

@@

The statistical analysis of the collected data will first focus on
examining the question if, on average, the usage of
\(\epsilon\)-lexicase selection will result in models that perform
significantly different than models that are evolved using tournament
selection both.

@@

\hypertarget{gentetic-programming-configuration}{%
\subsection{Gentetic Programming
Configuration}\label{gentetic-programming-configuration}}

@@

The primitive set consists of the terminals that are listed in table 3
and the functions that are listed in table 4. To avoid run time errors I
implemented protected version of the operators for division, natural
logarithm and square root (Koza, 1992, pp. 82--83). As suggested by Koza
(1992) I also included ephemeral constants to the set of terminals to
provide the evolutionary search the opportunity to explore and include
randomly generated constants.

@@

To gain further insight into the differences in generalization behavior
and to address the research question with a higher level of certainty,
the experiment I conducted should be repeated for different GP
configurations. Some exemplary parameters that could drastically
influence the results of this experiment include the population size,
the total number of generations, the inclusion of elitism or the
computation of \(\epsilon\) in lexicase selection. Other factors whose
influence on generalization behavior could be important include
different methods to compute an individuals fitness (e.g.~mean absolute
error or root mean squared error), testing different variations of the
genetic operators for mutation and crossover and experimenting with
different variations of the training:testing split size.

Another obvious weakness of my project is that the experiment is only
based on a single symbolic regression application, the prediction of the
heating load of buildings based on a relatively small data set. To gain
further trust in the obtained results, the experiment should be repeated
for other symbolic regression tasks and data sets. At the current state
it remains unknown if the generalization behavior of both selection
operators might be highly problem specific and dependent on the data set
used for training and testing.

>>>>

The overall performance of GP can depend strongly on the choice of its
underlying operators, one crucial component in this is the operator for
parent selection. Tournament Selection is a commonly used selection
operator in EC and is the most used operator in GP systems (Fang and Li,
2010, p. 181). A parent solution is selected by randomly sampling \(k\)
individuals from the current population into a tournament pool and then
the solution with the highest fitness score from the pool is selected
(Fang and Li, 2010, p. 182). Lexicase selection has been suggested as an
alternative to tournament selection that is not based on aggregating
fitness scores. It samples \(n\) test cases in random order and then
eliminates solutions from the selection pool on a per test case basis if
they are not performing on an elite level (Helmuth, Spector and
Matheson, 2015, p. 1).

@@

An important quality of all supervised machine learning applications,
including GP, is the ability to not only optimize performance for the
test cases a model is trained on but to also perform well on previously
unseen cases, this is referred to as generalization. In most real world
applications of symbolic regression only a small subset of labeled data
is available for training. The aim is to produce a model that not only
accurately predicts the provided training data but can also predict
previously unseen cases with high precision (Gonçalves, 2016, p. 6). A
model that is extensively optimized on training data, can become over
fitted to the data sample which may lead to a decrease in
generalization.

@@

Section 2 will give a brief summary and discussion of topic related
scientific research. Section 3 describes the design of the experiment as
well as GP specific implementation details. The statistical evaluation
of all relevant results and the summarized conclusions are given in
section 3 and 4. Section 5 addresses limitation and open questions of
this research project.

@@

Both articles argue that many of the problem domains that GP is commonly
used for are prevalent for problems that are modal or uncompromising in
nature. In comparison to other selection operators, populations that are
evolved using variations of the lexicase selection operator show a very
high degree of genetic diversity which might be a key contributor to the
improved performance (Helmuth, Spector and Matheson, 2015, p. 1) (La
Cava, Spector and Danai, 2016, p. 745). In theory, Lexicase based
selection should be more prone to select solutions that are specialist
(solutions that achieve high fitness on a smaller subset of unusual
training cases) than solutions that are generalist (solutions that
achieve high average fitness among all training cases).

@@

More recent research by La Cava, Spector and Danai (2016) suggests
\(\epsilon\)-lexicase selection as a modified selection operator that
can improve overall GP performance if applied to continuous-valued
symbolic regression tasks in comparison to tournament and standard
lexicase selection. La Cava, Spector and Danai (2016) argues that the
original concept behind lexicase selection as formulated by Spector
(2012) does not fit the requirements of real-world symbolic regression
tasks. The authors identifies the pass condition of regular lexicase
selection as the main problem if applied to symbolic regression:
Individual solutions can only be selected if they perform on an elite
level but with continuous-valued, and often noisy, data it is very
unlikely for two individuals to achieve an exact equal error for a given
test case. This stringent pass condition can result in filtering out too
many individuals during the selection process (La Cava, Spector and
Danai, 2016, p. 742). \(\epsilon\)-lexicase adresses this problem by
introducing an \(\epsilon\) parameter that specifies a range around the
elite error, individuals that perform inside this range pass the current
selection iteration. La Cava \emph{et al.} (2017) describes
\(\epsilon\)-lexicase selection as a more ``relaxed version of lexicase
selection''. Different methods to configure the \(\epsilon\) Parameter
have been explored, for the limited scope of this project I will focus
on the most promising implementation of \(\epsilon\)-lexicase selection
that is automatically adjusted on the basis of the median absolute
deviation of errors inside the selection pool (La Cava, Spector and
Danai, 2016, p. 742) (La Cava \emph{et al.}, 2017, p. 6).

@@

1. Sample:
  (a) Randomly sample k individuals from the current population
      into a tournament pool.
2. Select:
  (a) Compute the mean fitness for each individual inside the tournament pool
      based on all fitness cases.
  (b) Select individual from the tournament pool with the highest mean fitness score.
\end{verbatim}

@@

Generalization, the ability of a model to perform well on previously
unseen cases, is one of the fundamental goals in most real world machine
learning applications. O'Neill \emph{et al.} (2010) raises awareness to
the fact, that the topic of generalization in GP has not gotten the
attention that other machine learning-based areas attributed to it. A
similar statement has been proposed by Kushchu (2002) in his review of
research in generalization in GP. Both authors address the need for
additional research regarding the topic.

@@

An aspect that is suspected to negatively correlate with GP's
generalization is overall size and in particular bloating of the
resulting programs. Bloating can be described as a growth in the total
size of a program that does not improve its performance in any
meaningful way (Silva and Vanneschi, 2009, p. 1). It has been widely
suspected that GP configurations that produce very large programs also
have a higher tendency to specialize on difficult or unusual test cases
which could result in lower generalization (Wang, Wagner and Rondinelli,
2019, p. 268). The minimum description length principle (MDLP) describes
this phenomenon by claiming that less complex models are more likely to
perform better at generalizing than more complex models that achieve a
comparable fitness during the training phase (O'Neill \emph{et al.},
2010, p. 349). However MDLP should be considered with caution,
experiments by Silva and Vanneschi (2009) on the prediction of bio
availability for drugs demonstrate that GP based techniques can either
bloat without over fitting or over fit without bloating (Silva and
Vanneschi, 2009, p. 8).

@@

Since its inception, GP based symbolic regression has been widely used
in practical applications ranging over a wide field of disciplines. A
well known example of an industrial use-case has been published by
Jordaan \emph{et al.} (2004). The authors use GP based symbolic
regression to derive nonlinear functions that can be used to
significantly increase the robustness of industrial sensors. This
resulted in wide-spread adoption and a significant reduction in costs.

@@

Using two otherwise identical GP systems, one deploying tournament
selection and the other \(\epsilon\)-lexicase selection, the objective
is to find a computer program that best predicts the outcome variable
heating load \((Y1)\)\footnote{The symbolic regression is performed on
  one of the two provided outcome variables, the variable Cooling Load
  will be excluded.} of the buildings using a subset of the eight
building attributes \((X1,..,X8)\) for its input. The specific meaning
of all attributes inside the data set are described in table 1.

@@

The statistical analysis of the collected data will first focus on
examining the question if, on average, the usage of
\(\epsilon\)-lexicase selection will result in models that perform
significantly different than models that are evolved using tournament
selection:

@@

\hypertarget{genetic-programming-configuration}{%
\subsection{Genetic Programming
Configuration}\label{genetic-programming-configuration}}

@@

The primitive set consists of the terminals listed in table 3 and the
functions listed in table 4. To avoid run time errors I implemented
protected version of the operators for division, natural logarithm and
square root (Koza, 1992, pp. 82--83). As suggested by Koza (1992) I also
included ephemeral constants to the set of terminals to provide the
evolutionary search the opportunity to explore and include randomly
generated constants.

@@

To gain further insight into the differences in generalization behavior
and to address the research question with a higher level of certainty,
the experiment I conducted should be repeated for different GP
configurations. Some exemplary parameters that could drastically
influence the results of this experiment include the population size,
the total number of generations, the inclusion of elitism or the
computation of \(\epsilon\) in lexicase selection. Other factors whose
influence on generalization behavior could be important include
different methods to compute an individuals fitness (e.g.~mean absolute
error or root mean squared error), testing different variations of the
genetic operators for mutation and crossover and experimenting with
different variations of the training:testing data set split size.

Another obvious weakness of this project is that the experiment is only
based on a single symbolic regression application, the prediction of the
heating load of buildings based on a relatively small data set. To gain
further trust in the obtained results, the experiment should be repeated
for other symbolic regression tasks and data sets. It remains unknown if
the generalization behavior of both selection operators might be highly
problem specific and dependent on the data set used for training and
testing.
